---
title: "R Programming, Part 1"
author: "Kedge/Junia"
date: "4/22/2021"
output: html_document
---

In this first part of the course, we will understand the R language. First, we will explore the types of data that we can meet such as real numbers, characters, booleans...and understand the notion of variable.
Secondly, we will work on data structures, arrays (vectors) and matrices with the corresponding instructions.


# Data class and type
To begin,  we will talk about the numeric data class that represents numbers and digits in R.
## Numerics
```{r}
class(6.5)
class(6)
typeof(6.5)
as.integer(6.5)
typeof(as.integer(4))
#typeof(as.integer(6.5))
```


typeof (6.5) informs us that they are doubles.
If we want to transform this number into an int then we have to use as.integer (6.5) which returns us 6. working with integers or integrating does not take up much space in your machine's memory.
However, in this course, we will use the doubles.

## Characters
```{r}
class("hello")
typeof("hello")
typeof("v")
```

## Booleans 
Another class that we will discuss later is the Boolean class also called logical in R.
This is true or false for example
```{r}
44>23
44<23
```

## Variables
A variable will allow us to store information.
For example, when we need to get a file containing data, we will store it in a variable. So we can manipulate all of this data just by using a variable.

A variable will take several types of data as we saw earlier.
```{r}
my_variable = 4
my_variable
typeof(my_variable)
class(my_variable)
```

```{r}
my_caractere ="hello"
my_caractere
typeof(my_caractere)
class(my_caractere)
```


```{r}
my_booleen= TRUE

my_booleen
typeof(my_booleen)
bool= 4<2
bool
```

```{r}
Booleen = TRUE
```

# Arithmetic operations
What you need to know is that R works like a calculator
```{r}
4+6

var1 = 100 ## new variable named var1 which takes the value 100
var2 = 4 ## var2 takes the value 4
var1 + var2 ## addition between variables
var3 = var1 + var2 # assign the result of the addition to a new variable var3
var3 ## display of var3
var1 * var2 ## multiplication
var1 / var2 ## division
var1 %% var2 ## modulo operation (remainder of the division between var1 and var2)

```

# Simple data structure
## Vectors
A vector in R has the same definition as in linear algebra.
That is to say that it is a variable representing one or more variables.
If you are familiar with other programming languages, it's like an array of values.
The vector is represented by a variable where all the values are stored.
Earlier, we declared a variable "var1". This also represents a vector in R.

To create a vector we will use the function c for concatenation
```{r}
vec1= c(1,2,3)
vec1
typeof(vec1)
class(vec1)

```
If we want a character type vector

```{r}
c("a","b","c")
car1=c("a","b","c")
car1
```
You cannot mix data types in a vector.
The vector, like the variable, meets the typing criteria seen previously.
If we want to mix different types, for example, well let's experiment:


```{r, echo=TRUE}
vec2=c(3,"a",4.5)
vec2
typeof((vec2))
```
We note that by default, it is the type "character" which takes defined.

## Operations on vectors:
In this part, we are interested in arithmetic and other operations on vectors.
Other functions allow to generate vectors for example
```{r, echo=TRUE}
# with the seq function, two ways to write
my_vector1= seq(from=1, to=10)
my_vector1
my_vector2=seq(1:10)
my_vector2
```


```{r, echo=TRUE}
# with the rep function, two ways to write
my_vector3= rep(10, times = 10)
my_vector3
my_vector4 = rep(10,10)
my_vector4
```
To increase the values of the vector my_vector1 by 1

```{r, echo =TRUE}
my_vector1
my_vector1+1
my_vector1
```
To multiply the values of the vector my_vector1 by 2


```{r, echo=TRUE}
my_vector1
my_vector1*2

```


to do multiplication operations between two vectors
```{r, echo=TRUE}
my_vector1
my_vector2 = seq(from=6, to= 15)
my_vector2
my_vector1*my_vector2
```



Same for performing additions between vectors
```{r}
my_vector1
my_vector2
my_vector1+my_vector2
```

To perform a subtraction
```{r}
my_vector1
my_vector2
my_vector1 - my_vector2
```


To divide between vectors
```{r}
my_vector1
my_vector2
my_vector1 / my_vector2
```

To concatenate two vectors, just do the following operation
```{r}
my_vector1
my_vector3
my_vector5=c(my_vector1,my_vector3)
my_vector5

```


## The different functions for manipulating vectors:
In this section we will focus on existing and very useful functions in R which will allow us to manipulate vectors:
```{r}
length(my_vector1)  ## vector size

```
The length function takes a vector as argument.If we want for example to sort a vector from the smallest value to the largest value:

```{r}
my_vector6 = c(49, 24, 2, 87, 6, 10, 33, 56, 88, 17) ## creation of a new vector
sort(my_vector6) ## the sort function allows you to sort the values inside a vector.

```

If we want to create a new vector with the sorted values of the my_vector6 vector then
```{r}
mon_vecteurTrieC = sort(my_vector6)
mon_vecteurTrieC
```

Now if we want to sort descending, we just need to use the same output function with an option which is decreasing = TRUE:
```{r}
sort(my_vector6, decreasing= TRUE)

```
If we can also create a new variable or object to which we can assign this vector sorted in descending order:

```{r}
mon_vecteurTrieD=sort(my_vector6, decreasing = TRUE)
mon_vecteurTrieD
```

To see the ranks of each value in my array, we can use the rank function:

```{r}
my_vector6
rank(my_vector6) 
rank(mon_vecteurTrieC)
rank(mon_vecteurTrieD)
```
## Browse vectors
 
If, for example, we want to know the value at position 2 of the vector my_vector6
 
```{r}
my_vector6
my_vector6[2]
```
 
If we now want to display the first 3 values of the vector:
```{r}
my_vector6 [1: 3] ## the 1: 3 sequence indicates that I want the values from position 1 to position 3
my_vector6 [3: 7] ## displays values from position 3 to position 7
```

If we now wish to access positions 3, 5, 9 of the vector
```{r}
my_vector6
my_vector6[3]
my_vector6[c(3,5,9)]
```

If you want to retrieve all the values greater than 30
```{r}
my_vector6
superior = my_vector6> 30 ## creation of a new vector
superior ## here we can see that when it is a value greater than 30, it displays TRUE and when it is a value less than 30, it displays FALSE
my_vector6[superior] ## now, showing all values greater than 30
```

Ici, nous allons nous interesser à certaines fonctions qui permettent de faire certaines opérations sur les vecteurs: 
```{r}
my_vector6
sum (my_vector6) # sum of all vector values
mean (my_vector6) # mean of vector
min (my_vector6) # know the smallest value of the vector
max (my_vector6) # know the greatest value of the vector
median (my_vector6)
summary (my_vector6) # informs us of vector distributions

```

## Simple statistics on a vector
In this part, we are going to review the set of functions that we have seen previously by illustrating this with a simple exercise that speaks to everyone, that is to say the notes in a class.
Suppose we have math scores for a class of 25 students and we want to know the different statistics. In this case, it means knowing the average, the worst score, the highest score, those who are below the average, those who are above the average, etc.

For this, we will use a runif function, which will generate values randomly between a min and a max.
Then we want to round these values so that they are more integer type.

```{r}
grade_math = runif (25, min = 1, max = 20) # randomly generate 25 values between 1 and 20
grade_math # in this vector we can see that these are random values with lots of decimal places.
grade_math = round(grade_math, digits = 0) # here we are going to round the numbers generated by runif to any number after the decimal point with the round function.
grade_math
```


Now we are going to give names to the students
```{r}

names(grade_math) = c("Elise", "Leon","Sully", "Anouck","Baptiste", "Thibault", "Paul", "Zoé", "Zély", "Amandine", "Antoine", "Charles", "Lilou", "Chloé", "Daphnée", "Nicolas", "Maxcence", "Auguste", "Mathurin", "Sarah", "Lili", "Marion","Naomie","Raphael", "Flora")
grade_math

```



```{r}
typeof(grade_math)
class(grade_math)
```


## Edit vector values
```{r}
grade_math[3]= 14
grade_math["Baptiste"]= 13
names(grade_math[2])=c("Mikael")
grade_math
```


# Matrices
It is a table with two dimensions.
```{r}
vector1= sample(1:20, 12)
vector1
matrix(vector1,nrow=3, ncol = 4) 


```


```{r}
grade_class=runif(15, min=1, max=20)
grade_class=round(grade_class, digits = 0)
grade_class= matrix(grade_class, ncol = 3, nrow = 5)
grade_class
```

We can also create character type matrices
```{r}
ma_matrice2=matrix(c("a", "b", "c","d","e","f","g","h","i"), nrow = 3, ncol = 3)
ma_matrice2
```

In order to name the rows and columns of the matrices
```{r}
colnames(grade_class)= c("Biology", "Maths", "Physical")
rownames(grade_class)=c("Elise", "Leon","Sully", "Anouck","Baptiste")
grade_class
```

[LIGNE, COLONNE] toujours !

```{r}
grade_class [, "Biology"] ## displays the first column
grade_class [, 1] ## this instruction is equivalent to the previous one

```

```{r}
grade_class ["Sully",] ## displays the 3rd line
grade_class [3,] ## instruction equivalent to the previous one
```
If we want to see Sully's note in Physical, we have two ways:
```{r}
grade_class["Sully", "Physical"]
grade_class[3,3] ##instruction equivalent to the previous one

grade_class["Elise","Maths"]
grade_class[1,2] ## Elise's grade in Maths

```

To access Elise's Maths and Physical notes
```{r}
grade_class["Elise", c("Maths", "Physical")]
grade_class[1, c(2,3)]
grade_class["Elise", 2:3]
```

We can do the same actions at row level
For example see the notes of Anouck and Baptiste in Biology

```{r}
grade_class[c("Anouck","Baptiste"), "Biology"]

```

## Modify the elements of a matrix
```{r}
grade_class[c("Anouck","Baptiste"), "Biology"]= c(15, 17)
grade_class[c("Anouck","Baptiste"), "Biology"]
```

## Matrix operations
In this example, we will use the grade_class matrix and we will create a new matrix that we will name matrix_ope
```{r}
matrice_ope = c(rep(0.25, times = 5), rep(0.5, times = 5), rep(0.75, times = 5)) # the rep function allows you to repeat a number in times = 5 times
matrice_ope
matrice_ope=matrix(matrice_ope, nrow = 5, ncol = 3) # transformation of the vector into a matrix.
matrice_ope
```

## Addition of two matrices
```{r}
grade_class + matrice_ope
```

## Multiplication of two matrices
In R it is not a matrix calculation where it is a question of multiplying the first row of the first matrix by the first column etc. It is very basic, first column of the first matrix multiplied by first column of the second matrix.

```{r}
grade_class
matrice_ope
grade_class * matrice_ope
```

Likewise for the division
```{r}
grade_class
matrice_ope
grade_class/matrice_ope
```
##  The subtraction of two matrices
```{r}
grade_class
matrice_ope
grade_class-matrice_ope
```
 
## Some additional functions on matrices

```{r}
colSums(grade_class)
rowSums(grade_class)

colMeans(grade_class)
round(rowMeans(grade_class),digits = 2)
```

