---
title: "Shiny-COURSE-PART-2"
author: "Feryal WINDAL"
date: "11/5/2023"
output: html_document
---
# Your first shinydashboard 
## Introduction à Shinydashboard
Nous allons construire notre tout premier tableau de bord Shiny complet. Ce processus implique l'assemblage d'une interface utilisateur composée de trois parties principales : un en-tête, une barre latérale et un corps, tous combinés au sein de dashboardPage et attribués à la variable ui. L'étape suivante consiste à configurer la partie serveur où nous définissons les sorties et leur interaction avec les entrées des utilisateurs, semblable à un serveur qui prend les commandes et apporte les plats à la table. La fonction shinyApp est ensuite utilisée pour connecter l'UI et le serveur.

```{r}
header <- dashboardHeader(title="the first dashboard")
sidebar <- dashboardSidebar(...)
body <- dashboardBody(...)
ui <- dashboardPage(header,                    
                    sidebar, body)
server <-function(input, output){  ...}
shinyApp(ui, server)

```

## Configuration du Serveur
Dans la partie serveur d'une application Shiny, nous spécifions nos sorties et comment elles répondront aux entrées. Bien que les sorties puissent être définies directement dans la fonction serveur, la création de fonctions d'aide séparées les rend plus propres et organisées. Ces fonctions doivent être définies juste après les importations de bibliothèques dans notre script.

## Ajout d'un graphique 
Ajoutons un graphique à notre tableau de bord pour un tournoi de football mondial. Plus précisément, un graphique hexbin montrant les distributions de passes et de buts par une équipe. Nous utilisons la fonction geom_hex de ggplot2 et le définissons comme une sortie dans le serveur pour l'afficher sur le tableau de bord. Si nous voulons représenter des données pour une seconde équipe, nous pouvons le faire en introduisant une entrée pour sélectionner l'équipe et en rendant conditionnellement les graphiques respectifs.
Création d'Objets ggplot dans le Serveur


# Exercise

```{r}
library(tidyverse)
soccer <-read_csv("soccer18-19.csv")
```


```{r}

head(soccer)
```

```{r}
p<-soccer %>%  
  ggplot(aes(x=`FTHG`,y=`HTR`)) +  
  geom_hex(bins=10) + 
  theme_classic()

p
```
#Header and side bar settings
Suppose you are working for an apparel company. Since a global soccer tournament has just concluded, your company wants to produce football apparel. To catch the attention of potential customers, the company needs to be in touch with the fan base.

You have been tasked to create a dashboard to help the company make better design decisions, based on the following:

Some details for each match
Total number of goals scored for each team
Total number of yellow/red cards given to each team


# The body, and completing the UI
You have set up the header and sidebar Now, you can set up the body.

This is a crucial part of the wireframing process, since it will help you decide where to position your outputs and inputs in their desired positions within the main body of the dashboard.

In this exercise, you will place several outputs and inputs in the dashboard. In doing so, you will obtain all three components of the user interface, which will be put together in dashboardPage().

Set the second tab name to "statstab", and place a tabBox() containing three sub tabs called "Goals", "Yellow cards" and "Red cards".

```{r}
header <- dashboardHeader(title = "Soccer tournament")

sidebar <- dashboardSidebar(
  sidebarMenu(
    menuItem("Matches", tabName = "matchtab", icon = icon("futbol")), 
    menuItem("General statistics", tabName = "statstab", icon = icon("chart-line")) 
  )
)

body <- dashboardBody(tabItems(
  tabItem(tabName = "matchtab", 
  fluidRow(selectInput("match", "Match number", choices = 1:nrow(soccer))) 
         ),
  # See that the second tab should contain three sub-tabs
  ___(tabName = ___, 
          ___(tabPanel("Goals"), 
                 ___("Yellow cards"),
                 ___(___))
         )
))
ui <- dashboardPage(header, sidebar, body)
server <- function(input, output){}
shinyApp(ui, server)
```


Add three infoBoxOutput()s, called "daytime", "venue" and "grp", to the first row of the first page.
To the second row of the first page, add four valueBoxOutput()s called "team1", "score1", "team2" and "score2".

```{r}
header <- dashboardHeader(title = "Soccer tournament")

sidebar <- dashboardSidebar(
  sidebarMenu(
    menuItem("Matches", tabName = "matchtab", icon = icon("futbol")), 
    menuItem("General statistics", tabName = "statstab", icon = icon("chart-line")) 
  )
)

body <- dashboardBody(tabItems(
  tabItem(tabName = "matchtab", 
  fluidRow(selectInput("match", "Match number", choices = 1:nrow(soccer)),
  # Add three info boxes called "daytime", "venue" and "grp"
           infoBoxOutput("daytime"), ___("venue"), ___),
  # Add four value boxes called "team1", "score1", "team2" and "score2"
  fluidRow(valueBoxOutput("team1", width = 3), ___("score1", width = 3), 
           ___(___, width = 3), ___(___, width = 3)),
  fluidRow(plotOutput("histogram"))),
  tabItem(tabName = "statstab", 
          tabBox(tabPanel("Goals", plotOutput("goals", height = "700px")),
                 tabPanel("Yellow cards", plotOutput("yellow", height = "700px")),
                 tabPanel("Red cards", plotOutput("red"))))
))
ui <- dashboardPage(header, sidebar, body)
server <- function(input, output){}
shinyApp(ui, server)
```

Now that we have the header, side bar and body of the user interface (UI) stored as header, sidebar and body, respectively, place them in dashboardPage() and store the UI as ui.


```{r}
library(shiny)
library(shinydashboard)

header <- dashboardHeader(title = "Soccer tournament")

sidebar <- dashboardSidebar(
  sidebarMenu(
    menuItem("Matches", tabName = "matchtab", icon = icon("futbol")),
    menuItem("General statistics", tabName = "statstab", icon = icon("chart-line"))
  )
)

body <- dashboardBody(tabItems(
  tabItem(tabName = "matchtab", 
  fluidRow(selectInput("match", "Match number", choices = 1:nrow(soccer)),
  # Add three info boxes called "daytime", "venue" and "grp"
           infoBoxOutput("daytime"), infoBoxOutput("venue"), infoBoxOutput("grp")),
  # Add four value boxes called "team1", "score1", "team2" and "score2"
  fluidRow(valueBoxOutput("team1", width = 3), valueBoxOutput("score1", width = 3), 
           valueBoxOutput("team2", width = 3), valueBoxOutput("score2", width = 3)),
  fluidRow(plotOutput("histogram"))),
  tabItem(tabName = "statstab", 
          tabBox(tabPanel("Goals", plotOutput("goals", height = "700px")),
                 tabPanel("Yellow cards", plotOutput("yellow", height = "700px")),
                 tabPanel("Red cards", plotOutput("red"))))
))
ui <- dashboardPage(header, sidebar, body)
server <- function(input, output){}
shinyApp(ui, server)
```

# Building your first dashboard
The data has been stored as a data frame called soccer. There are 3 predefined plots:

goals_plot: The number of goals scored by each team
yellow_plot, red_plot: The number of yellow/red cards given to each team
Some custom helper functions are used here:

daytime_fn(): Gets the day & time for a particular match
venue_fn(): Gets the venue for a particular match
grp_fn(): Gets the group of the matchup
team1_fn(), team2_fn(): The names of team 1/2
score1_fn(), score2_fn(): The number of goals scored by team 1/2

```{r}
body <- dashboardBody(tabItems(
  tabItem(tabName = "matchtab", 
  fluidRow(selectInput("match", "Match number", choices = 1:nrow(soccer)),
           infoBoxOutput("daytime"), infoBoxOutput("venue"), infoBoxOutput("grp")),
  fluidRow(valueBoxOutput("team1", width = 3), valueBoxOutput("score1", width = 3), 
           valueBoxOutput("team2", width = 3), valueBoxOutput("score2", width = 3)),
  fluidRow(plotOutput("histogram"))),
  tabItem(tabName = "statstab", 
          tabBox(tabPanel("Goals", plotOutput("goals", height = "700px")),
                 tabPanel("Yellow cards", plotOutput("yellow", height = "700px")),
                 tabPanel("Red cards", plotOutput("red"))))
))

ui <- dashboardPage(header, sidebar, body)

server <- function(input, output) {
  # Fill in outputs in first to third rows of the first page
  output$daytime <- renderInfoBox(infoBox("Day and time", 
                                          daytime_fn(input$match), 
                                          icon = icon("calendar"), 
                                          color = "green"))
  output$venue <- ___(infoBox("Venue", 
                                        venue_fn(input$match), 
                                        icon = icon("map"), 
                                        color = "green"))
  output$___ <- ___(___("Group", 
                                      grp_fn(input$match), 
                                      color = "green"))
  output$team1 <- ___(valueBox("Team 1", team1_fn(input$match), color = "blue"))
  output$score1 <- ___(___("# of goals", score1_fn(input$match), color = "blue"))
  output$___ <- ___(valueBox("Team 2", team2_fn(input$match), color = "red")) 
  ___ <- ___(valueBox("# of goals", score2_fn(input$match), color = "red"))
  output$histogram <- renderPlot(plot_histogram(input$match))
  # Fill in outputs in the second page
  output$goals <- ___(goals_plot)
  output$___ <- ___(yellow_plot)
  ___ <- ___(red_plot)
}

# Put the UI and server together
ui <- dashboardPage(header, sidebar, body)
server <- function(input, output){}
shinyApp(ui, server)
```


To simplify your work, I have defined a set of functions that allow you to manipulate the dataset and simplify the coding. In its raw state, this dataset does not easily yield information

```{r}
daytime_fn <- function(match_number) {
  soccer %>%
    filter(row_number() == match_number) %>%
    pull(Date)
}

grp_fn <- function(match_number) {
  soccer %>%
    filter(row_number() == match_number) %>%
    pull(Div)
}

team1_fn <- function(match_number) {
  soccer %>%
    filter(row_number() == match_number) %>%
    pull(HomeTeam)
}

team2_fn <- function(match_number) {
  soccer %>%
    filter(row_number() == match_number) %>%
    pull(AwayTeam)
}

score1_fn <- function(match_number) {
  soccer %>%
    filter(row_number() == match_number) %>%
    pull(FTHG)
}

score2_fn <- function(match_number) {
  soccer %>%
    filter(row_number() == match_number) %>%
    pull(FTAG)
}

venue_fn <- function(match_number) {
  # Comme il n'y a pas de colonne 'Venue', nous utilisons 'HomeTeam' comme lieu
  soccer %>%
    filter(row_number() == match_number) %>%
    pull(HomeTeam)
}

# Cette fonction va créer un histogramme des buts à domicile pour tous les matchs
plot_histogram <- function() {
  hist(soccer$FTHG, main = "Histogram of Full Time Home Goals", xlab = "Full Time Home Goals", breaks = 10)
}


```



```{r}
library(shiny)
library(shinydashboard)
header <- dashboardHeader(title = "Soccer tournament")

sidebar <- dashboardSidebar(
  sidebarMenu(
    menuItem("Matches", tabName = "matchtab", icon = icon("futbol")),
    menuItem("General statistics", tabName = "statstab", icon = icon("chart-line"))
  )
)

body <- dashboardBody(tabItems(
  tabItem(tabName = "matchtab", 
  fluidRow(selectInput("match", "Match number", choices = 1:nrow(soccer)),
           infoBoxOutput("daytime"), infoBoxOutput("venue"), infoBoxOutput("grp")),
  fluidRow(valueBoxOutput("team1", width = 3), valueBoxOutput("score1", width = 3), 
           valueBoxOutput("team2", width = 3), valueBoxOutput("score2", width = 3)),
  fluidRow(plotOutput("histogram"))),
  tabItem(tabName = "statstab", 
          tabBox(tabPanel("Goals", plotOutput("goals", height = "700px")),
                 tabPanel("Yellow cards", plotOutput("yellow", height = "700px")),
                 tabPanel("Red cards", plotOutput("red"))))
))

ui <- dashboardPage(header, sidebar, body)

server <- function(input, output) {
  # Fill in outputs in first to third rows of the first page
  output$daytime <- renderInfoBox({
    infoBox("Day and time", daytime_fn(input$match), icon = icon("calendar"), color = "green")
  })
  output$venue <- renderInfoBox({
    infoBox("Venue", venue_fn(input$match), icon = icon("map"), color = "green")
  })
  output$grp <- renderInfoBox({
    infoBox("Group", grp_fn(input$match), color = "green")
  })
  output$team1 <- renderValueBox({
    valueBox("Team 1", team1_fn(input$match), color = "blue")
  })
  output$score1 <- renderValueBox({
    valueBox("# of goals", score1_fn(input$match), color = "blue")
  })
  output$team2 <- renderValueBox({
    valueBox("Team 2", team2_fn(input$match), color = "red")
  })
  output$score2 <- renderValueBox({
    valueBox("# of goals", score2_fn(input$match), color = "red")
  })
  
  # Pour l'histogramme, nous n'utilisons pas input$match car il affiche les buts pour tous les matchs
  output$histogram <- renderPlot({
    plot_histogram()
  })
  
  # Assumez que vous avez les fonctions `goals_plot`, `yellow_plot`, et `red_plot` définies quelque part pour rendre ces graphiques
  output$goals <- renderPlot({ goals_plot() })
  output$yellow <- renderPlot({ yellow_plot() })
  output$red <- renderPlot({ red_plot() })
}


# Put the UI and server together
shinyApp(ui, server)
```

